---
title: "Uncerain Arctic main analysis script",
autors: "Angrea Bryndum-Buchholz, Julia Mason, Juliano Palacios-Abrantes"
output: html_document
date: "2023-09-19"
---


Maps:
- Add lat/lon axis labels (if possible)
- Increase margin to include all LMEs fully
- Modify color bar labels to a full number range
- Label each LME with a number, corresponding with the Y-axis labels in bargraphs
- Label land masses




```{r results = "hide", warnings = F, message = F, eval = T}

# Load needed functions
source("../functions/load_libraries.R")
library(MyFunctions)

load_libs(c(
  "tidyverse",
  "ncdf4",
  "sf",
  "viridis",
  "wesanderson",
  "rlme"
)
)

# library(reticulate)
# library(tidyverse)
# library(metR)
# library(lubridate)
# library(raster)
# library(sf)

# Data path
data_path <- "~/Library/CloudStorage/OneDrive-UBC/Data/FishMip/cmip6_runs/csv/"

```

# Methods

## Load Data

### LME grid

Data from [Protection Of The Marine Environment: PAME](https://pame.is/projects/ecosystem-approach/arctic-large-marine-ecosystems-lme-s). Large Marine Ecosystems (LMEs) are regions of ocean space of 200,000 km² or greater, that encompass coastal areas from river basins and estuaries to the outer margins of a continental shelf or the seaward extent of a predominant coastal current. LMEs are defined by ecological criteria, including bathymetry, hydrography, productivity, and tropically linked populations. The LME concept for ecosystem-based management has 5-module approach focused on productivity, fish and fisheries, pollution and ecosystem health, socioeconomics, and governance.

```{r lme_grid, eval = T}

# Get Polar Large Marine Ecoregions
lme_sf <- my_sf("Arctic_LME") %>% 
  #Fix names for plots
  mutate(lme = paste0(str_remove(lme, "LME"),"(",id,")"))
crs <- st_crs(lme_sf)

unique(lme_sf$lme)

# ggplot(lme_sf) +
#   geom_sf(aes(fill = lme))
  # theme(legend.position = "")

# Get world map data and crop it to the polar region
world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf") %>% 
  st_transform(world, crs = crs)

fishmip_grid <- read_csv(paste0("~/Library/CloudStorage/OneDrive-UBC/Data/FishMip/fishmip_grid_1deg.csv"))

# For some reason DBPM and ZOOMs for IPSL are in another grid account 
# extra_grid <- read_csv(paste0("~/Library/CloudStorage/OneDrive-UBC/Data/FishMip/zoom_dbpm_ipsl_grid_1deg.csv"))

# Rasterize

# All of fishmip
fishmip_sf <- st_as_sf(fishmip_grid,
                      coords = c("lon","lat"),
                      crs = 4326
                      )

# Join FishMip Meow grid
fishmip_lme_df <- st_join(fishmip_sf,lme_sf %>% st_transform(crs = 4326)) %>% 
  as.data.frame() %>% 
  select(rowid,lme)

unique(fishmip_lme_df$lme)

```

### FishMip data

```{r load_fishmip_data, eval = T, echo = F, message = F}

ssp <- "585"

# FishMip cmip6 data
historical_list <- list.files(data_path,full.names = T, pattern = "historical")

hist_df <- bind_rows(
  lapply(historical_list, read_csv)
)

model_list <- list.files(data_path,full.names = T, pattern = ssp)

future_df <- bind_rows(
  lapply(model_list, read_csv)
  ) %>% 
  mutate(ssp = as.character(ssp))


all_models <- bind_rows(hist_df,future_df)

# unique(all_models$mem)
# unique(all_models$ssp)

# Joint all data needed

models_df <- all_models %>% 
  left_join(fishmip_grid,
            by =c("lat", "lon")
            ) %>% 
  left_join(fishmip_lme_df,
            by = "rowid") %>% 
  group_by(lme,esm,year,mem,ssp) %>% 
  summarise(tcb = sum(tcb, na.rm = T)) %>% 
  filter(!is.na(lme))

unique(models_df$mem)
unique(models_df$ssp)
unique(models_df$lme)

```

## Proportion change

```{r}

historic_data <- models_df %>% 
  filter(ssp == "historical",
         year > 1995) %>% 
  group_by(lme,esm,mem) %>% 
  summarise(mean_hist = mean(tcb, na.rm = T),
            sd_hst = sd(tcb, na.rm = T)
            )


delta_df <- models_df  %>% 
  filter(ssp != "historical",
         year > 2030 & year < 2049) %>% 
  group_by(lme,esm,mem) %>% 
  summarise(period_mean = mean(tcb,na.rm= T)) %>% 
  left_join(historic_data,
            by = join_by("lme", "esm", "mem")
  ) %>% 
  # mutate(change = period_mean-mean_hist) %>%
  mutate(change = ((period_mean-mean_hist)/mean_hist)*100,
         direction = case_when(
           change > 0 ~ "Increase",
           change < 0 ~ "Decrease",
           TRUE ~ "No Change"
           )
         ) %>% 
  group_by(lme,esm) %>% 
  summarise(
    mean_mem = mean(change,na.rm = T),
    sd_mem = sd(change,na.rm = T),
    n_mems = length(unique(mem)),
    mems = paste(unique(mem),collapse = ";"),
    n_direction = max(sum(direction == "Increase", na.rm = TRUE),sum(direction == "Decrease", na.rm = TRUE))/n_mems) %>% 
  filter(n_mems>1) %>% 
  pivot_longer(
    cols = c(mean_mem,sd_mem,n_direction)
    )


# max(delta_df$n_mems)
# unique(models_df$mem)
# unique(models_df$lme)

```

# Results

# All ESMs and all MEMs

```{r}

agg_delta <- delta_df %>% 
  group_by(lme,name) %>% 
  summarise(
    value = mean(value,na.rm = T)
  )
```

### Bar plot on mean and sd

Bar plots:
- Modify color bar labels to a full number range ()
- change colour of color bar (to not be the same as on the map)
- fix typos in colour bar title and LME names
- Label each LME with a number, corresponding with the LME labels in map
- delete "LME" after LME names
- add (%) after TCB on X-axis

```{r}

agg_delta %>%
  spread(name,value) %>% 
  # filter(!is.na(lme)) %>% 
  mutate(flag = ifelse(sd_mem > abs(mean_mem),"Larger","Smaller")) %>% 
  # View()
  ggplot() +
  geom_bar( 
    aes(
      y = reorder(lme,mean_mem),
      x = mean_mem,
      fill = flag),
    stat = "identity"
  ) +
    geom_errorbar(
      aes(
        y = lme,
        xmin = mean_mem - sd_mem, 
         xmax = mean_mem + sd_mem#,
        # color = flag
        ),
      color = "black"
      ) +
  labs(x = "Change in TCB (%)",
         y = "Large Marine Ecosystem (LME)") +
  scale_fill_manual(values = wesanderson::wes_palette("GrandBudapest2", type = "discrete"),
                    breaks = seq(0,1,0.2)
                    ) +
  scale_color_manual(values = wesanderson::wes_palette("GrandBudapest2", type = "discrete"),
                     breaks = seq(0,1,0.2)
                     ) +
  my_ggtheme_p(
    ax_tl_s = 12,
    ax_tx_s = 9,
    leg_tl_s = 12,
    leg_tx_s = 9
    )

ggsave(plot = last_plot(),
       paste0("../figures/lme_agg_bar_flag_",ssp,".png"),
       width = 10)
  

```


### Bar plot on direction

```{r}

agg_delta %>%
  spread(name,value) %>% 
  filter(!is.na(lme)) %>% 
  ggplot() +
  geom_bar( 
    aes(
      y = reorder(lme,mean_mem),
      x = mean_mem,
      fill = n_direction),
    stat = "identity"
  ) +
  geom_errorbar(
    aes(
      y = lme,
      xmin = mean_mem - sd_mem, 
      xmax = mean_mem + sd_mem
    ),
    color = "grey30"
  ) +
  labs(x = "Change in TCB (%)",
       y = "Large Marine Ecosystem (LME)") +
  scale_fill_viridis("Direction Index",
                     option = "A",
                     direction = -1,
                     alpha = 0.8,
                     # limits = c(0,1),
                     # breaks = seq(0,1,0.2)
                     ) +
  my_ggtheme_p(
    ax_tl_s = 12,
    ax_tx_s = 9,
    leg_pos = "right",
    leg_tl_s = 12,
    leg_tx_s = 9
    )

ggsave(plot = last_plot(),
       paste0("../figures/lme_agg_bar_plot_dir_",ssp,".png"),
       width = 10)
  
```


### Maps

Plots maps for average, s.d. and direction change

```{r}

for(i in 1:length(unique(agg_delta$name))){
  
  variable <- unique(agg_delta$name)[i]
  
  if(variable == "n_direction"){
    leg_title = "Direction index"
  }
  
  if(variable == "sd_mem"){
    leg_title = "Variation\n(%)"
  }
  
  if(variable == "mean_mem"){
    leg_title = "Change in TCB\n(%)"
  }
  
  
  figure_name <- paste0("../figures/lme_",variable,"_agg_map_",ssp,".png")
  
  lme_sf %>% 
    left_join(agg_delta %>% filter(name == variable)) %>%
    ggplot() +
    geom_sf( 
      aes(
        fill = value),
      color = "grey30"
    ) +
    geom_sf(data = world,aes(), fill = "antiquewhite")+
    geom_sf_label(aes(label = id, fill = value), color = "black") +
    scale_color_viridis(leg_title, option = "D",direction = -1) +
    scale_fill_viridis(leg_title, option = "D",direction = -1) +
    scale_x_continuous("Longitude", breaks = c(-50,-30,0,30)) +
    scale_y_continuous("Latiitude", breaks = c(35,40,45,50)) +
    coord_sf(crs = crs, 
             xlim = c(-4000000, 2300000), 
             ylim = c(-3100000, 4100000)) +
    MyFunctions::my_ggtheme_m(map_type = "reg",
                              leg_pos = "right"
    ) +
    theme(
      # axis.text.y = element_blank(),
          legend.title.align = 0.5
    )
  
  ggsave(plot = last_plot(),
         figure_name,
         width = 6,
         height = 6)
  
}

```


## By ESM

### Bar plot on mean and sd

```{r}

delta_df %>%
  spread(name,value) %>% 
  mutate(flag = ifelse(sd_mem > abs(mean_mem),"Larger","Smaller")) %>% 
  ggplot() +
  geom_bar( 
    aes(
      y = reorder(lme,mean_mem),
      x = mean_mem,
      fill = flag),
    stat = "identity"
  ) +
    geom_errorbar(
      aes(
        y = lme,
        xmin = mean_mem - sd_mem, 
         xmax = mean_mem + sd_mem
        ),
      color = "black"
      ) +
  labs(x = "Change in TCB",
         y = "lme") +
  scale_fill_manual(values = wesanderson::wes_palette("GrandBudapest2", type = "discrete")) +
  scale_color_manual(values = wesanderson::wes_palette("GrandBudapest2", type = "discrete")) +
  facet_wrap(~esm)

ggsave(plot = last_plot(),
       "../figures/lme_bar_plot_flag_85.png",
       width = 10)
  

```

### Bar plot on direction

```{r}

delta_df %>%
  spread(name,value) %>% 
  filter(!is.na(lme)) %>% 
  ggplot() +
  geom_bar( 
    aes(
      y = reorder(lme,mean_mem),
      x = mean_mem,
      fill = n_direction),
    stat = "identity"
  ) +
  geom_errorbar(
    aes(
      y = lme,
      xmin = mean_mem - sd_mem, 
      xmax = mean_mem + sd_mem#,
      # color = flag
    ),
    color = "grey30"
  ) +
  labs(x = "Change in TCB",
       y = "LME") +
  scale_fill_viridis("Diection index",
                     option = "D",
                     direction = -1,
                     alpha = 0.8) + 
  facet_wrap(~esm)

ggsave(plot = last_plot(),
       "../figures/lme_bar_plot_dir_85.png",
       width = 10,
       height = 7)
  
```

### Maps

Plots maps for average, sd and direction change

```{r}

for(i in 1:length(unique(delta_df$name))){
  
  variable <- unique(delta_df$name)[i]
  
  if(variable == "n_direction"){
    leg_title = "Direction index"
  }
  
  if(variable == "sd_mem"){
    leg_title = "Variation\n(%)"
  }
  
  if(variable == "mean_mem"){
    leg_title = "Mean change\n(%)"
  }
  
  
  figure_name <- paste0("../figures/lme_",variable,"_map_85.png")
  
  lme_sf %>% 
    left_join(delta_df %>% filter(name == variable)) %>%
    ggplot() +
    geom_sf( 
      aes(
        fill = value),
      color = "grey30"
    ) +
    geom_sf(data = world,aes(), fill = "antiquewhite")+
    scale_color_viridis(leg_title, option = "D",direction = -1) +
    scale_fill_viridis(leg_title, option = "D",direction = -1) +
    facet_wrap(~esm) +
    coord_sf(crs = crs, 
             xlim = c(-2300000, 2300000), 
             ylim = c(-2300000, 2300000)) +
    MyFunctions::my_ggtheme_m(map_type = "reg",
                              leg_pos = "right"
    ) +
    theme(axis.text.y = element_blank(),
          legend.title.align = 0.5
    )
  
  ggsave(plot = last_plot(),
         figure_name,
         width = 8,
         height = 4)
  
}

```

# Old code 

## Grid level analysis

### MEOW grid

```{r meow_grid, eval = F}


meow <- MyFunctions::my_sf("MEOW")

# ggplot(meow) +
#   geom_sf(aes(fill = ecoregion)) +
#   theme(legend.position = "") +
#   coord_sf(ylim = c(66,90))


fishmip_grid <- read_csv(paste0("~/Library/CloudStorage/OneDrive-UBC/Data/FishMip/fishmip_grid_1deg.csv"))

# For some reason DBPM and ZOOMs for IPSL are in another grid account 
# extra_grid <- read_csv(paste0("~/Library/CloudStorage/OneDrive-UBC/Data/FishMip/zoom_dbpm_ipsl_grid_1deg.csv"))

# Rasterize

# All of fishmip
fishmip_sf <- st_as_sf(fishmip_grid,
                      coords = c("lon","lat"),
                      crs = 4326
                      )
# Special kids
# extra_sf <- st_as_sf(extra_grid,
#                       coords = c("lon","lat"),
#                       crs = 4326
#                       )

# Join FishMip Meow grid
fishmip_meow_df <- st_join(fishmip_sf,meow) %>% 
  as.data.frame() %>% 
  select(rowid,ecoregion,province,realm)




# Join extra kids grid
# extra_meow_df <- st_join(meow,extra_sf) %>% 
#   as.data.frame() %>% 
  # select(rowid,ecoregion,province,realm)


```

### Proportion change

```{r, eval = F}

historic_data <- all_models %>% 
  filter(ssp == "historical",
         year > 1995) %>% 
  group_by(lat,lon,esm,mem) %>% 
  summarise(mean_hist = mean(tcb, na.rm = T),
            sd_hst = sd(tcb, na.rm = T)
            )


delta_df <- models_df  %>% 
  filter(ssp != "historical",
         year > 2030 & year < 2049) %>% 
  group_by(lat,lon,esm,mem) %>% 
  summarise(period_mean = mean(tcb,na.rm= T)) %>% 
  left_join(historic_data,
            by = join_by("lon", "lat", "esm", "mem")
  ) %>% 
  # mutate(change = period_mean-mean_hist) %>%
  mutate(change = ((period_mean-mean_hist)/mean_hist)*100) %>%
  group_by(lat,lon,esm) %>% 
  summarise(
    mean_mem = mean(change,na.rm = T),
    sd_mem = sd(change,na.rm = T),
    n_mems = length(unique(mem)),
    mems = paste(unique(mem),collapse = ";")
  ) %>% 
  filter(n_mems>1) %>% 
  pivot_longer(
    cols = c(mean_mem,sd_mem)
    )

# 
max(delta_df$n_mems)

unique(models_df$mem)


```


### Grid Cell

```{r, eval = F}

ggplot() +
  geom_tile(data = mean_data, 
            aes(
              x = lon,
              y = lat,
              fill = value_plot,
              color = value_plot)
            ) +
  geom_polygon(data = map_data("world"),
               aes(long, lat, group = group),
               fill = "grey50",
               color = "black") +
  coord_map("orthographic",ylim = c(70,90)) +
  # scale_colour_gradient2("Delta Change") +
  # scale_fill_gradient2("Delta Change")+
  scale_color_viridis(option = "D") +
  scale_fill_viridis(option = "D") +
  facet_wrap(~esm)

```

### Using countour 
https://www.rdocumentation.org/packages/metR/versions/0.14.0/topics/geom_contour_fill

```{r, eval = F}

ggplot() +
  geom_contour_filled(data = mean_data,
     aes(
       x = lon,
       y = lat,
       z = value_plot,
       fill = stat(level)
     ),
     bins = 5
  ) +
  geom_polygon(data = map_data("world"),
               aes(long, lat, group = group),
               fill = "grey50",
               color = "black") +
  coord_map("orthographic",ylim = c(70,90)) +
  # scale_colour_gradient2("Delta Change") +
  # scale_fill_gradient2("Delta Change")+
  facet_wrap(~esm)
  

```


```{r, eval = F}


ggplot() +
  geom_tile(data = sd_data, 
            aes(
              x = lon,
              y =lat,
              # fill = value_plot,
              # color = value_plot
              fill = log10(value_plot),
              color = log10(value_plot)
            )
            ) +
  geom_polygon(data = map_data("world"),
               aes(long, lat, group = group),
               fill = "grey50",
               color = "black") +
  coord_map("orthographic",ylim = c(70,90)) +
  scale_color_viridis_b() +
  scale_fill_viridis_b() +
  facet_wrap(~esm)
  

```


```{r, eval = F}

ggplot() +
  geom_contour_filled(data = sd_data, 
     aes(
       x = lon,
       y = lat,
       z = value,
       fill = stat(level)
     )
  ) +
  geom_polygon(data = map_data("world"),
               aes(long, lat, group = group),
               fill = "grey50",
               color = "black") +
  coord_map("orthographic",ylim = c(70,90)) +
  facet_wrap(~esm)
  

```

